---
title: "Student Performance"
author: "Didier , Bertrand , Reda"
date: "1/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Student Performance (Mini project) : 


### I - Préparation des données :

> Chargement des library : 

```{r}
library(dplyr)
library(tidyverse)
library(readxl)
library(broom)
library(pastecs)
```

> Importation des donnee des Etudiants :

```{r}
data <- read_excel("MP-5RBIG.xlsx")
```

### 1 Verification des valeurs manquantes : 

```{r}
sapply(data,function(x) sum(is.na(x)))
```

### 2 Ajout de la variable binaire finalResult :

```{r}
newDf <- data %>%
  mutate(finalResutl=ifelse(G3>=10,"Pass","Fail"))
```

### 3 Ajout de la varialbe AcademicGreade : 

```{r}
newDf <- newDf %>%
    mutate(academicGrade = case_when(
      G3>=16 & G3 <=20 ~ "A",
      G3>=14 & G3 <=15 ~ "B",
      G3>=12 & G3 <=13 ~ "C",
      G3>=10& G3 <=11 ~ "D",
      G3>=0 & G3 <=9 ~ "E",
    ))
```


### 4 Afficher la structure des donnee : 

```{r}
str(newDf)
```


***

## II Etude des variables Categorielles :

### 1 Etude des principales varialbe catégorielles : 

School , sex , Pstatus , Studytime  , faliure  , health , internet   , FinalResult , AcademicGrade 

* Etude de la variable School :
 + Frequence & Proportion:

```{r}
  #Frequence:
  school.table <- table(newDf$school)
  #Proportion : 
  prop.table(school.table)
```
 + Diagrame qui represente la frequence des etudiant par ecole :

```{r}
ggplot(newDf , aes(school)) +
  geom_bar(fill = "cornflowerblue", 
           color="black") +
  labs(x = "School", 
       y = "Frequency", 
       title = "Student by School")
```

 + Chi square Test good of fit :
 
 H0 : les valeur sont distribuer comme suive : 80%:Gp , 20:Ms
 
```{r}
chisq.test(school.table , p=c(0.9 , 0.1))
```
 ***
 * Etude de la variable sex :
  + Frequence & Proportion:

```{r}
  #Frequence:
  sex.table <- table(newDf$sex)
  #Proportion : 
  prop.table(sex.table)
```
 + Diagrame qui represente la frequence des etudiant par sex :

```{r}
ggplot(newDf , aes(sex)) +
  geom_bar(fill = "deepskyblue4", 
           color="black") +
  labs(x = "sex", 
       y = "Frequency", 
       title = "Student by sex")
```

 + Chi square Test good of fit :
 
 H0 : les valeur sont distribuer comme suive : 80%:Gp , 20:Ms
 
```{r}
chisq.test(sex.table)
```

***
 * Etude de la variable faliure :
  + Frequence & Proportion:

```{r}
  #Frequence:
  failures.table <- table(newDf$failures)
  #Proportion : 
  prop.table(failures.table)
```
 + Diagrame qui represente la frequence des etudiant par sex :

```{r}
ggplot(newDf , aes(failures )) +
  geom_bar(fill = "deepskyblue4", 
           color="black") +
  labs(x = "failures ", 
       y = "Frequency", 
       title = "Student by failures ")
```

 + Chi square Test good of fit :
 
 H0 : les valeur sont distribuer comme suive : 80%:Gp , 20:Ms
 
```{r}
chisq.test(failures.table)
```

***
 * Etude de la variable internet :
  + Frequence & Proportion:

```{r}
  #Frequence:
  internet.table <- table(newDf$internet)
  #Proportion : 
  prop.table(internet.table)
```
 + Diagrame qui represente la frequence des etudiant par sex :

```{r}
ggplot(newDf , aes(internet )) +
  geom_bar(fill = "deepskyblue4", 
           color="black") +
  labs(x = "internet ", 
       y = "Frequency", 
       title = "Student by internet ")
```

 + Chi square Test good of fit :
 
 H0 : les valeur sont distribuer comme suive : 80%:Gp , 20:Ms
 
```{r}
chisq.test(internet.table)
```

***
 * Etude de la variable finalResutl :
  + Frequence & Proportion:

```{r}
  #Frequence:
  finalResutl.table <- table(newDf$finalResutl)
  #Proportion : 
  prop.table(finalResutl.table)
```
 + Diagrame qui represente la frequence des etudiant par sex :

```{r}
ggplot(newDf , aes(finalResutl )) +
  geom_bar(fill = "deepskyblue4", 
           color="black") +
  labs(x = "finalResutl ", 
       y = "Frequency", 
       title = "Student par Resultat final ")
```

 + Chi square Test good of fit :
 
 H0 : les valeur sont distribuer comme suive : 80%:Gp , 20:Ms
 
```{r}
chisq.test(finalResutl.table)
```

***
 * Etude de la variable academicGrade :
  + Frequence & Proportion:

```{r}
  #Frequence:
  academicGrade.table <- table(newDf$academicGrade)
  #Proportion : 
  prop.table(academicGrade.table)
```
 + Diagrame qui represente la frequence des etudiant par sex :

```{r}
ggplot(newDf , aes(academicGrade )) +
  geom_bar(fill = "deepskyblue4", 
           color="black") +
  labs(x = "academicGrade ", 
       y = "Frequency", 
       title = "Student par Notes academic ")
```

 + Chi square Test good of fit :
 
 H0 : les valeur sont distribuer comme suive : 80%:Gp , 20:Ms
 
```{r}
chisq.test(academicGrade.table)
```

***

### 2 Représentation des associations entre des paires de variables catégorielles : 

+ Representing the association between sex ~ final result

```{r}
ggplot(newDf, 
       aes(x = finalResutl, 
           fill = sex)) + 
  geom_bar(position = "stack") +
  scale_fill_brewer(palette = "Set2") +
  geom_text(
    aes(label=after_stat(count)),
    stat = 'count',
    position = position_stack(vjust = 0.5)
  )
```

+ Representing the association between Studytime ~ Academic Grade 

```{r}
ggplot(newDf, 
       aes(x = academicGrade, 
           fill = factor(studytime , 
                         levels= c("1" , "2" ,"3", "4"),
                         labels = c("Less then 2h" , "Between 2h & 5h" , "Between 5h & 10h" , "Between 4h & 10h")
                         ))) + 
  geom_bar(position = "fill") +
  labs(y = "Proportion", 
       fill = "Study Time",
       x = "Academic Grade ",
       title = "Academic Grades by  studytime") +
  theme_minimal()
```

+ Representing the association between Internet ~ finalResult

```{r}
internetByfinalR <- newDf %>%
  group_by(internet, finalResutl) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

internetByfinalR
```
```{r}
ggplot(internetByfinalR, 
       aes(x = finalResutl,
           y = pct,
           fill = internet)) + 
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(breaks = seq(0, 1, .2), 
                     label = waiver()) +
  geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "iorikeet") +
  labs(y = "Percent", 
       fill = "Internet",
       x = "Final Result",
       title = "Final Result by Internet") +
  theme_minimal()
```

+ Representing the association between Father job  ~ final results

```{r}
ggplot(newDf, aes(x=Fjob , group=finalResutl,fill=finalResutl)) + 
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Set1") +
  geom_text(
    aes(label=after_stat(count)),
    stat = 'count',
     position = position_dodge(0.9)
  )
```

+ Representing the association between final Result  ~ studytime


```{r}
failureBystudytime <- newDf %>%
  group_by(failures, studytime) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

failureBystudytime

ggplot(failureBystudytime,aes(failures , y=pct , fill=factor(studytime , 
                         levels= c("1" , "2" ,"3", "4"),
                         labels = c("Less then 2h" , "Between 2h & 5h" , "Between 5h & 10h" , "Between 4h & 10h")
                         ))) + 
  geom_bar(stat = "identity",
           position = "fill") +
   scale_y_continuous(breaks = seq(0, 1, .2), 
                     label = waiver()) +
  geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Set3") +
  labs(y = "Percent", 
       fill = "Study Time",
       x = "Failures",
       title = "Failures  by Studytime") +
  theme_minimal()
```


```{r}
ggplot(newDf, aes(x=higher, group=finalResutl,fill=finalResutl)) + 
  geom_bar() +
  geom_text(
    aes(label=after_stat(count)),
    stat = 'count',
    position = position_stack(vjust = 0.5))
  
```

***

### 3 Study associations between pairs of categorical variables with Chi-squared tests : 

The Hypotheses are :

H0 : the variables are independent, there is no relationship between the two categorical variables. Knowing the value of one variable does not help to predict the value of the other variable

H1: the variables are dependent, there is a relationship between the two categorical variables. Knowing the value of one variable helps to predict the value of the other variable




  * Study association between Sex and final result 
  
    + calculating 2-Way Frequency Table
    
```{r}
asc_sex_finalR <- table(newDf$sex , newDf$finalResutl)
```
    
   + Performing the Chi-Square Test of Independence : 
   
```{r}
tidy(chisq.test(asc_sex_finalR))
```
we see that the p-value is less than the significance level of 5%. Like any other statistical test, if the 
p-value is less than the significance level, we can reject the null hypothesis.
    
  * Study association between Internet and final result 
  
    + calculating 2-Way Frequency Table
    
```{r}
asc_internet_finalR <- table(newDf$internet , newDf$finalResutl)
```
    
  + Performing the Chi-Square Teest of Independence :
    
```{r}
tidy(chisq.test(asc_internet_finalR))
```
    
    
   * Study association between StudyTime and failures 
  
    + calculating 2-Way Frequency Table
```{r}
asc_studyT_failure <- table(newDf$studytime , newDf$failures)
```
    
   + Performing the Chi-Square Teest of Independence : 
    
```{r}
tidy(chisq.test(asc_studyT_failure))
```
  

   * Study association between studytime and final result 
  
    + calculating 2-Way Frequency Table
```{r}
asc_studyT_finalR <- table(newDf$studytime , newDf$finalResutl)
```

   + Performing the Chi-Square Test of Independence : 
```{r}
tidy(chisq.test(asc_internet_finalR))
```

    
  * Study association between Sex and final result 
  
    + calculating 2-Way Frequency Table
```{r}
asc_famrel_finalR <- table(newDf$famrel  , newDf$finalResutl)
```
    
    
   + Performing the Chi-Square Test of Independence : 
    
```{r}
tidy(chisq.test(asc_famrel_finalR))
```
  

***

### III - Etude des variables quantitatives :

### 1 Étudiez individuellement les principales variables quantitatives : 

  * statistiques descriptives :
  
```{r}
# Selecting quantitatives variales :
q_var <- newDf %>% 
  select(absences , age , G1 , G2 , G3)
summary(q_var)
```
```{r}
options(scipen=100)
options(digits=2)
stat.desc(q_var)
```
  
  * Ploting somme graphique that describe the destribution of a single quantitative variable : 
  
   + Age distribution : 
   
```{r}
library(scales)  
ggplot(newDf , aes(x=age , y= ..count.. / sum(..count..))) + 
  geom_histogram(fill = "cornflowerblue", 
                 color = "white") +
  labs(title = "Student by Age",
       x = "Age" , y="Percent") +
  scale_y_continuous(labels = percent , limits = c(0, 0.3))
```
   
   + G1 (first period grade in Math) distribution :
   
```{r}
ggplot(newDf , aes(x=G1)) + 
  geom_histogram(bins=30 , fill = "coral1", 
                 color = "white") +
  labs(title = "Student by G1",
       x = "G1" , y="Count")
```
   
  + G2 (Second period grade in Math) distribution :
  
```{r}
ggplot(newDf , aes(x=G2)) + 
  geom_histogram(bins=30 , fill = "coral2", 
                 color = "white") +
  labs(title = "Student by G2",
       x = "G2" , y="Count")
```

+ G2 (Second period grade in Math) distribution :

```{r}
ggplot(newDf , aes(x=G3)) + 
  geom_histogram(bins=30 , fill = "coral3", 
                 color = "white") +
  labs(title = "Student by G3",
       x = "G3" , y="Count")
```



